---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# StratifiedSampling package

In this R package, different functions are implemented for selecting samples . 

* If the population of interest is stratified. Different functions are implemented, for more details see <https://doi.org/10.1007/s42081-021-00134-y>.
* If two datasets are available for statistical matching. A method based on optimal transport is implemented, for more details see <https://arxiv.org/abs/2105.08379>.
* If you are interested in the Sequential Spatially Balanced method. <http://arxiv.org/abs/2112.01164>

The package contains also some useful functions. Look at the manual of the package for more information. 

## Installation

### CRAN version

```
install.packages("StratifiedSampling")
```

### Latest version 

You can install the latest version of the package `StratifiedSampling` with the following command:

``` r
# install.packages("devtools")
devtools::install_github("Rjauslin/StratifiedSampling")
```

## Optimal transport matching

The package proposes a method to do statistical matching using optimal transport and balanced sampling. For more details see Raphaël Jauslin and Yves Tillé (2021) <https://arxiv.org/abs/2105.08379>. A complete example on how to use the package to make an optimal statistical transport match can be found in the following vignette:

```
vignette("ot_matching", package = "StratifiedSampling")
```

## Sequential spatially balanced sampling

The package proposes a method to select a well-spread sample balanced on some auxiliary variables. For more details see Raphaël Jauslin and Yves Tillé (2022) <https://arxiv.org/abs/2112.01164>. A complete example on how to use the different functions to select a well-spread and balanced sample can be found in the following vignette:
```
vignette("sequential_balanced", package = "StratifiedSampling")
```


## Simple example on stratified population

Integrating a stratified structure in the population in a sampling design can considerably reduce the variance of the Horvitz-Thompson estimator. We propose in this package different methods to handle the selection of a balanced sample in stratified population. For more details see Raphaël Jauslin, Esther Eustache and Yves Tillé (2021) <https://doi.org/10.1007/s42081-021-00134-y>.



This basic example shows you how to set up a stratified sampling design. The example is done on the `swissmunicipalities` dataset from the package `sampling`.

```{r}
library(sampling)
library(StratifiedSampling)

data(swissmunicipalities)
swiss <- swissmunicipalities
X <- cbind(swiss$HApoly,
        swiss$Surfacesbois,
        swiss$P00BMTOT,
        swiss$P00BWTOT,
        swiss$POPTOT,
        swiss$Pop020,
        swiss$Pop2040,
        swiss$Pop4065,
        swiss$Pop65P,
        swiss$H00PTOT )

X <- X[order(swiss$REG),]
strata <- swiss$REG[order(swiss$REG)]
```

Strata are NUTS region of the Switzerland. Inclusion probabilities `pik` is set up equal within strata and such that the sum of the inclusion probabilities within strata is equal to 80.

```{r}
pik <- sampling::inclusionprobastrata(strata,rep(80,7))
```


It remains to use the function `stratifiedcube()`.

```{r}
s <- stratifiedcube(X,strata,pik)
```



We can check that we have correctly selected the sample. It is balanced and have the right number of units selected in each stratum.

```{r}
head(s)

sum(s)
t(X/pik)%*%s
t(X/pik)%*%pik

Xcat <- disj(strata)

t(Xcat)%*%s
t(Xcat)%*%pik

```

