---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# Stratified Sampling

  
## Build 
[![Build Status](https://travis-ci.org/RJauslin/WaveSampling.svg?branch=master)](https://travis-ci.org/RJauslin/WaveSampling)

## Installation

<!-- ### CRAN version -->

<!-- ``` -->
<!-- install.packages("WaveSampling") -->
<!-- ``` -->

### Latest version 

You can install the latest version of the package `StratifiedSampling` with the following command:

``` r
# install.packages("devtools")
devtools::install_github("Rjauslin/StratifiedSampling")
```

## Simple example

This basic example shows you how to solve a common problem. The example is done on the `swissmunicipalities` dataset from the package `sampling`.

```{r}
library(sampling)
library(StratifiedSampling)

data(swissmunicipalities)
swiss <- swissmunicipalities
X <- cbind(swiss$HApoly,
        swiss$Surfacesbois,
        swiss$P00BMTOT,
        swiss$P00BWTOT,
        swiss$POPTOT,
        swiss$Pop020,
        swiss$Pop2040,
        swiss$Pop4065,
        swiss$Pop65P,
        swiss$H00PTOT )

```

Inclusion probabilities `pik` is set up all equal with the function `rep()`.

```{r}
pik <- inclusionprobabilities(swiss$POPTOT,400)
```


It only remains to use the function `wave()`,

```{r}
s <- stratifiedcube(X,swiss$REG,pik)
```

