---
title: "Sequential balanced sampling"
author: "Raphaël Jauslin"
date: "`r Sys.Date()`"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
bibliography: ref.bib
vignette: >
  %\VignetteIndexEntry{Sequential balanced sampling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
# library(StratifiedSampling)

```


## Introduction


In this vignette we are going to explain how to use the function `balseq` to select a balanced sample. 

## Load data
We will use the dataset `belgianmunicipalties` from the package `sampling`. The dataset does not contains any spatial coordinates. Luckily, a `GEOjson` is directly available from the \url{"https://hub.arcgis.com/datasets/esribeluxdata::belgium-municipalities-1/about"}. We can then transform it into a `sf` object and calculate the centroid of the municipalities. We will use these centroid as spatial coordinates to spread the sample over the space.

```{r}
# to load data
library(sampling)
library(geojsonio)

# plot library
library(ggplot2) 
library(viridis)

# to transform geojson into sf and get centroid
library(rgeos) 
library(sf)
library(rmapshaper)



data("belgianmunicipalities")

# load geojson directly from the url
belg  <- geojson_read("https://opendata.arcgis.com/datasets/9589d9e5e5904f1ea8d245b54f51b4fd_0.geojson",what = "sp")

# simplify the variable and transform it into a sf object
belg <- rmapshaper::ms_simplify(input = belg,keep = 0.01) %>%
  st_as_sf()

coord <- gCentroid(as(belg, "Spatial"), byid = TRUE)

# concatenated file
Belgium <- cbind(belg,belgianmunicipalities,coord)
head(Belgium)

```

A `sf` object can be directly plotted using the function `geom_sf` from the package `ggplot2`.

```{r, fig.width= 10,fig.height=10,out.width = '100%'}
p <- ggplot()+
  geom_sf(data = Belgium,aes(fill = averageincome),size = 0.1)+
  scale_fill_viridis_c(option = "G")
p
```

## Inclusion probabilites

Firstly, we define the variable of interest, the auxiliary variables and the inclusion probabilities. We set up here the inclusion probabilities equal with sum equal to 50. i.e. the sample will contain 50 units.


```{r}

N <- nrow(Belgium) # population total
n <- 50 # sample size

# variable of interest
y <- belgianmunicipalities$averageincome

# auxiliary variables
Xaux <- cbind(belgianmunicipalities$Tot04,
              belgianmunicipalities$Women04,
              belgianmunicipalities$TaxableIncome,
              belgianmunicipalities$Diffmen,
              belgianmunicipalities$Diffwom)


# inclusion probabilities
pik <- rep(n/N,N)

Xaux <- cbind(pik,Xaux) # add pik to fixed sample size
Xspread <- coord

```


## Balanced sampling

We can now load the package `StratifiedSampling` and use the function `balseq` on the variables that we just defined. We can check that our method is selecting effectively a balanced sample by observing if the auxiliary totals are respected compared to a maximal entropy design (also know as conditional Poisson sampling). 

```{r}

library(StratifiedSampling)

s <- balseq(pik,Xaux)
s_cube <- samplecube(Xaux,pik,comment = FALSE)
s_srs <- srswor(n,N)


TOT <- colSums(Xaux)
EST1 <- colSums(Xaux[s,]/pik[s])
EST2 <- colSums(Xaux[s_cube == 1,]/pik[s_cube == 1])
EST3 <- colSums(Xaux[s_srs == 1,]/pik[s_srs == 1])


100* (EST1 - TOT)/TOT
100* (EST2 - TOT)/TOT
100* (EST3 - TOT)/TOT

```


## Spread sampling

To add the spatial component, we can simply add the spatial coordinates as a matrix to the argument `Xspread` of the function. Here `coord` is an output of the function `gCentroid` which is by construction an `S4` object. To get the `data.frame` that are encapsulated inside, we simply use the `@coords` operator.

```{r,fig.width=10,fig.height=10,out.width='100%'}

s <-  balseq(pik,
             Xaux,
             Xspread = as.matrix(Xspread@coords))
p <- p + 
  geom_point(data = Belgium,aes(x = x,y = y),shape = 1,alpha = 0.9)+
  geom_point(data = Belgium[s,],aes(x = x,y = y),colour = "red")+
  scale_fill_viridis_c(option = "G")
p

```

